%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[1]{Publications}{Scientific Publications}
\chapter*{Scientific Publications}
Some of the ideas, images and data exposed in this thesis have been previously published in the next references:
\bigskip

\begin{itemize}
\item \person{Pablo García-Sánchez, J. González, Pedro A. Castillo, Maribel García Arenas, Juan Julián Merelo Guervós} \emph{Service oriented evolutionary algorithms}.  Soft Comput. 17(6): 1059--1075 (2013).

\item \person{Pablo Garc\'ia-S\'anchez, Jes\'us Gonz\'alez, Antonio Miguel Mora, Maribel Garc\'ia Arenas, Pedro A. Castillo, Carlos Fernandes and Juan Juli\'an Merelo}. \emph{Population size adaptation in distributed evolutionary algorithms on heterogeneous clusters}. Under review in Applied Soft Computing.


\item \person{Pablo García-Sánchez, Maria I. García Arenas, Antonio Miguel Mora, Pedro A. Castillo, Carlos Fernandes, Paloma de las Cuevas, Gustavo Romero, Jesús González, Juan Julián Merelo Guervós} \emph{ Developing services in a service oriented architecture for evolutionary algorithms}. In Proceeding of the fifteenth annual conference companion on Genetic and evolutionary computation conference companion. ACM, 2013. p: 1341-1348.

\item \person{Pablo García-Sánchez} \emph{ A service oriented evolutionary architecture: applications and results}. In Proceeding of the fifteenth annual conference companion on Genetic and evolutionary computation conference companion. ACM, 2013. p: 1663--1666.

\item \person{Pablo García-Sánchez, J. González, Pedro A. Castillo, Juan Julián Merelo Guervós, Antonio Miguel Mora, Juan Luís Jiménez Laredo, Maribel García Arenas} \emph{ A Distributed Service Oriented Framework for Metaheuristics Using a Public Standard}. In Proceeding of Nature Inspired Cooperative Strategies for Optimization. Studies in Computational Intelligence. Springer, 2010. p: 211--222.

\item \person{P. García-Sánchez, J. J. Merelo, D. Calandria, A. B. Pelegrina, R. Morcillo, F. Palacio and R. H. García-Ortega} \emph{Testing the Differences of using RGB and HSV Histograms during Evolution in Evolutionary Art } In proceedings of the 5th International Joint Conference on Computational Intelligence. 2013. p. 168-174.

\item \person{P. García-Sánchez, A. Fernández-Ares, A. M. Mora, P. A. Castillo, J. González and J.J. Merelo} \emph{Tree depth influence in Genetic Programming for generation of competitive agents for RTS games}. Applications of Evolutionary Computation, EvoApplicatons 2010: EvoCOMPLEX, EvoGAMES, EvoIASP, EvoINTELLIGENCE, EvoNUM, and EvoSTOC, Proceedings. Springer, 2013 Lecture Notes in Computer Science (to appear).

%\item \person{José M. Palomares, Jesús González, Eduardo Ros y Alberto Prieto.} \emph{General Logarithmic Image Processing Convolution}. IEEE Transactions on Image Processing, 15(11):3602--3608, 2006. \\
%{\color{Maroon}DOI: 10.1109/tip.2006.881967}. Índice de Impacto: $2.715$\\
%\item \person{José M. Palomares, Jesús González y Eduardo Ros.} \emph{Designing a fast convolution under the {LIP} paradigm applied to edge detection} En Sameer Singh \etAl (editor): \emph{Proceedings of the III ICAPR 2005, Bath (Reino Unido)} publicado en \emph{Lecture Notes on Computer Sciences,} LNCS 3687(3):560--569, 2005. {\color{Maroon}DOI: 10.1007/11552499\_62}.\\
%Índice de Impacto: $0.402$\\
%\item \person{José M. Palomares, Jesús González y Eduardo Ros.} \emph{Detección de bordes en imágenes con sombras mediante LIP--Canny} En las Actas del I Simposio de Reconocimiento de Formas y Análisis de Imágenes: \emph{Conferencia AERFAI 2005, Granada (España)}:71--76, 2005. {\color{Maroon} ISBN:84--9732--445--5}\\
\end{itemize}
